---
title       : Introduction to Piketty's Capital
subtitle    : Thomas Piketty, Capital in the 21st Century (Harvard University Press 2014)
author      : Patrick Toche
framework   : io2012  # {minimal, io2012, html5slides, shower, dzslides, ...}
mode        : selfcontained  # {selfcontained, standalone, draft}
widgets     : mathjax   # {mathjax, quiz, bootstrap}
ext_widgets : {rCharts: [libraries/nvd3]}  # libraries/morris, libraries/highcharts, libraries/rickshaw, libraries/dimple
url         : {lib: "../../libraries", assets: "../../assets"}
lecnum      : "00"
--- {class: cover-slide}

```{r 'preamble', message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, comment = NA, cache = FALSE}
require(knitr)
library(png)   # to display external png and control dimensions
library(grid)  # to display external png and control dimensions
opts_chunk$set(echo = FALSE, cache = FALSE)
source('../../shared/shared.R')
```

---

### Introduction

- This set of slides surveys selected topics from *Capital in the Twenty-First Century*, a book written by economist Thomas Piketty, published in English in 2014 to great acclaim.
- All source files for this course are available for download by anyone without restrictions at https://github.com/ptoche/piketty
- The full course is expected to be completed by April 2015.
- The intended audience for these notes is any curious person with or without training in economics.
- The book deals with technical matters but its streamlined presentation of technical issues makes it highly readable.

---

### Thomas Piketty

<figure class = "centered">  
<img src = "../../images/Piketty_BloombergBusinessWeek_2014.jpg" alt = "Thomas Thomas Piketty (born 1971).">
<figcaption class = 'figcaption'>Piketty's Capital took America by storm. 
</figcaption>  
</figure> 

---

### 21世紀の資本

- 私達は所得の分配について何を知っているか。
- 私達は富の分配について何を知っているか。
- 所得と富の長期的な進化についてどんな予測ができるか。
- 民間の資本蓄積の力がより一層少ない人々に富を集中させることは避けられないのか。
- 成長、競争および技術的な進歩を均衡させる力は不平等を減らし階級間の衝突を減らすのに充分強くなり得るか。
- この導入部は不平等の非常に簡単な歴史を提示し、本書の主題について描写する。

--- 

### Adam Smith

<figure class = "centered">  
<img src = "../../images/Adam_Smith.jpg" alt = "Adam Smith (1723-1790).">
<figcaption class = 'figcaption'>Adam Smith (1723-1790).
</figcaption>  
</figure> 

--- 

### アダム・スミスと貧困

- 1776年にアダム・スミスは彼の研究を *『国富論』* で出版した。
- アダム・スミスは「最初の経済学者」だと考えられている。彼が生きた社会では、貧困、飢餓、若年の死が一般的であった。この混沌の中で、彼は経済を奇跡、「見えざる手」の仕業以外の何物でもないと考えた。
- スミスにとって、豊かな社会は不平等であり、遅れた社会の不平等がより少ない限り、その社会は貧しい社会である。
- 資本蓄積は資本の限界生産性を下げ、労働の限界生産性を上げ、平等を減らすことに寄与するとスミスは考えた。
- 社会が発展するにつれて、平等と豊かさが達成されるとスミスは期待した。

---

### Thomas Robert Malthus

<figure class = "centered">  
<img src = "../../images/Thomas_Robert_Malthus.jpg" alt = "Thomas Robert Malthus (1766-1834).">
<figcaption class = 'figcaption'>Reverend Thomas Malthus (1766-1834).
</figcaption>  
</figure> 

---

### マルサスと人口過剰の問題
 
- 1798年、トマス・マルサスが *人口論* を出版した。彼の主な懸念は人口過剰であった。
- フランス王国の人口は1700年の2000万人から1780年までに3000万人に増加した。この非常に急速な人口増加により、1789年の革命以前の数十年で、農業の賃金上昇が停滞し地代が上昇した。
- 国家がより豊かになると、人口が増え、典型的な貧しい家庭の資源は減少したので、貧しい者はより貧しくなった。この不幸な逆説がマルサスを悩ませた。
- 人口過剰、混乱、不幸を防ぐために、貧しい者の出産は減らされなければならず、その実現のために福祉の支援は削減されなければならないとマルサスは主張した。

---

### David Ricardo

<figure class = "centered">  
<img src = "../../images/David_Ricardo_by_Thomas_Phillips.jpg" alt = "David Ricardo (1772-1823).">
<figcaption class = 'figcaption'>David Ricardo (1772-1823), Portrait by Thomas Phillips.
</figcaption>  
</figure> 

---

### リカードと希少性の原理

- 1817年、リカードは *『経済学および課税の原理』* を出版した。彼の主な研究対象は、土地の価格と地代の長期的な進化であった。
- リカードは資本主義が内部に矛盾を持つと考えた。
- 人口と生産が継続的に増加するとき、土地は他の財に比べてますます希少になる傾向がある。すると需要と供給の法則が意味するのは、地主に支払われる地代と同様に、土地の価格が継続的に上昇するということである。従って国民所得に占める地主の取り分が増加し、その分人口の残りの取り分が減る。
- リカードにとって、政治的に受容できる解決策は地代に対する課税を継続的に増加させることであった。

---

### リカードと希少性の原理

- リカードの予測は間違いであると証明された。国民所得に占める農業の割合が減少し、農地の相対的な価値が低下したのである。
- リカードは技術革新とそれが希少性の原理を相殺する役割を持つことの重要性を予測していなかった。
- 希少性の原理は今も重要である。農地の価格を都市部の不動産に置き換え、1970年から2010年の価格の趨勢に基づいて2010年から2050年について推測すると富の格差がどうなるかを考察せよ。
- 需要と供給の仕組みは結局のところ極端な趨勢に対して反対の効果を持つ。ある財の供給が減り価格が高すぎる場合、その財に対する需要は減少し、結果として価格は下がる。

---

### Karl Marx

<figure class = "centered">  
<img src = "../../images/Karl_Marx_1875.jpg" alt = "Karl Marx (1818-1883).">
<figcaption class = 'figcaption'>Karl Marx (1818-1883), Portrait dated 1875.
</figcaption>  
</figure> 

---

### マルクスと資本論

- 1867年、マルクスが *『資本論』* の第1部を出版した。他の2部は彼の死後 *『共産党宣言』* (1848年)の共著者フリードリヒ・エンゲルスによって出版された。
- 問題は、農業従事者が増加する人口を養うことができるかどうかとか、土地の価格が上昇するかどうかではなかった。
- 19世紀の前半には、賃金が非常に低い水準で停滞し、前の世紀より低いことさえあった。
- 19世紀前半のイギリスとフランスでは、資本所得の割合がかなり増加した。
- 産業革命と農村地帯からの大規模な人口の移動に伴い、労働者が都市部のスラムにごった返した。その時貧困は都市部で極端な豊かさと並んで見られるようになった。

---

### マルクスと利益率の低下

- マルクスは、資本主義者が際限なく富を蓄積する傾向があることと、富がより一層少ない人々に集中する傾向があることを予言した。
- マルクスは過度な資本の蓄積が利益率を低下させ産業資本主義を崩壊させると予言した。
- マルクスは利益率が低下する傾向を「政治経済の最も重要な法則」であると考えた。利益率が低下するにつれて、資本主義者はより多くの剰余を搾取しようとし、労働者の賃金は停滞するか減少する。
- マルクスの予言は間違いであると証明された。19世紀の最後の3分の1で、賃金はついに上昇し始めた。
- ヨーロッパの最も遅れた国、ロシアで共産主義革命が起こったが、先進国では起こらなかった。

---

### Simon Kuznets and the Kuznets Curve

- In 1953, Kuznets published *Shares of Upper Income Groups in Income and Savings*, a study of the United States over 1913–1948. 
- Kuznets noted a sharp reduction in the upper deciles of income between 1913 and 1948. In 1913, the top 10% of US earners received 45–50% of annual national income. In 1948, they earned 30–35%.
- The 'Kuznets curve' interpretation is that inequality rises in the early stages of industrialization &mdash; falls later. An explanation is due to W. Arthur Lewis (1915-1991). In early stages, 'unlimited' supply of labour from rural areas allows urban areas to expand without upward pressure on wages. Wages rise once rural-urban migration stops.
- In truth, however, the compression of high incomes was largely accidental, caused by the Great Depression and World War II.

---

### Piketty's Capital

- While Marx's vision was a nightmare, the Kuznets curve is a fairy tale.
- Piketty seeks to ground his theories on facts. He surveys economic data spanning several centuries and continents.
- Piketty's theory places the data within a simple, universal accounting framework and focuses on the very long run.
- Piketty's vision of the future draws from history and does not speculate about future technologies or future societies.
- This 'lack of imagination' leads Piketty to search for fiscal solutions to the economic and political problem of inequality.
- Piketty advocates high taxes on high incomes and wealth transmission.

---

### Forces of Convergence & Divergence

- **Convergence**
    - The main force is the diffusion of knowledge.
    - depends on education policies, access to skills, institutions.
- **Divergence**
    - `1.` Top earners earn hundreds of times average earnings. 
    - `2.` Wealth concentration tends to occur when economic growth is smaller than the rate of return on wealth.
    - We have become accustomed to 1, thanks to the work of Thomas Piketty and Emmanuel Saez, and many others.
    - We are now discovering 2 thanks to Piketty's book and the dire warnings of Larry Summers, Paul Krugman, and others.

---

### Forces of Convergence & Divergence

- Figures I.1 and I.2 show two basic patterns of divergence. 
    - Figure I.1 depicts income inequality in the United States.
    - Figure I.2 depict the capital/income ratio in 3 European countries. 
- Both graphs depict "U-shaped curves": A period of decreasing inequality followed by one of increasing inequality. 
- But the phenomena depicted in the figures involve distinct economic, social, and political processes. 

---

### Forces of Convergence & Divergence

- Figure I.1 shows the share of the upper decile of the income hierarchy in US national income from 1910 to 2010. 
- The top decile claimed as much as 45–50% of national income in the 1910s–1920s, fell to 30–35% by the end of the 1940s, rose again in the 1980s, and by 2000 reached about 45–50%.
- Figure I.2 shows the total value of private wealth (in real estate, financial assets, and professional capital, net of debt) in Britain, France and Germany, expressed in years of national income, for 1870–2010. 
- The total amount of private wealth was about 6 or 7 years of national income in the 19th c., fell sharply to 2 or 3 years, in response to the shocks of the period 1914–1945, started to rise steadily after 1950, and is up to about 5 years in 2010.

---

### Income Inequality in the United States

- The spectacular increase in inequality largely reflects an unprecedented explosion of very elevated incomes from labor, caused by a separation of the top managers of large firms from the rest of the population.
- Did the skills and productivity of these top managers rise suddenly in relation to those of other workers? 
- Do we live in a 'winner takes all' society?
- Top managers suddenly acquired the power to set their own remuneration. The 'winner grabs all' society!
- The tendency is less marked in other wealthy countries, but the trend is in the same direction.

---

### The capital-income ratio

- The rise of capital/income ratios over the past few decades is in large part a return to a regime of slow growth. 
- If the rate of return on capital remains above the growth rate for an extended period of time, then wealth becomes more concentrated.
- In slowly growing economies, past wealth matters more.
- The logic is captured by an inequality:
$$ 
r > g
$$
where r stands for the average annual rate of return on capital, including profits, dividends, interest, rents, and other income from capital, and g stands for the rate of growth of the economy, that is, the annual proportional change in total income.

<!---

### Income inequality in the United States

<figure class = "captioned">  
<img class = "captioned" src = "../../figures/Figure_0_1.png" alt = "Figure I.1. Income inequality in the United States, 1910-2012.">  
<figcaption class = 'figcaption'>Figure I.1. The top decile share in U.S. national income dropped from 45-50% in the 1910s-1920s to less than 35% in the 1950s (this is the 1950-1960 fall documented by Kuznets); it then rose from less than 35% in the 1970s to 45-50% in the 2000s-2010s.
</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

-->

---

### Income inequality in the United States

```{r 'Figure_0_1_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_0_1.Rda") 
# round data for rChart tooltip display
df_0_1$value <- round(df_0_1$value, 3)
n <- nPlot(data = df_0_1, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(.3, .5))
n$yAxis(axisLabel = 'Share of top decile in national income')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(showLegend = FALSE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_0_1.html', standalone = TRUE)
``` 
<iframe src = 'figures/Figure_0_1.html' alt = "Figure I.1. Income inequality in the United States, 1910-2012.">
</iframe><icaption class = 'icaption'>Figure I.1. The top decile share in U.S. national income dropped from 45-50% in the 1910s-1920s to less than 35% in the 1950s (this is the 1950-1960 fall documented by Kuznets); it then rose from less than 35% in the 1970s to 45-50% in the 2000s-2010s.</icaption> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

<!--- Comment out this slide

### The capital-income ratio in Europe

<figure class = "captioned">  
<img class = "captioned" src = "../../figures/Figure_0_2.png" alt = "Figure I.2. The growth of world population, 1700-2012.">  
<figcaption class = 'figcaption'>Figure I.2. Aggregate private wealth was worth about 6-7 years of national income in Europe in 1910, between 2 and 3 years in 1950, and between 4 and 6 years in 2010.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

-->

---

### The capital-income ratio in Europe

```{r 'Figure_0_2_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_0_2.Rda") 
# round data for rChart tooltip display
df_0_2$value <- round(df_0_2$value, 3)
n <- nPlot(data = df_0_2, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(1.5, 7.5))
n$yAxis(axisLabel = 'Market value of private capital (% national income)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  }
</style>')
n$save('figures/Figure_0_2.html', standalone = TRUE)
``` 
<iframe src = 'figures/Figure_0_2.html' alt = "Figure I.2. The capital-income ratio in Europe, 1870-2012.">
</iframe><icaption class = 'icaption'>Figure I.2. Aggregate private wealth was worth about 6-7 years of national income in Europe in 1910, between 2 and 3 years in 1950, and between 4 and 6 years in 2010.</icaption> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

